\documentclass[../main.tex]{subfiles}

\begin{document}


\section{Related Work}
\justify

A structured search across Google Scholar, IEEE Xplore, and Semantic Scholar over five topics -- bark-beetle detection, obstacle avoidance, drone sensors, reinforcement learning for UAVs, and sim-to-real transfer -- limited to 2019--2025, narrowed about 450 results to 25 after duplicate, title, and abstract screening; three foundational papers \cite{Schulman2017,Pinto2017,Hrabar2012} were added by hand, giving 28 references. Studies are grouped into classical single-camera navigation, deep reinforcement learning (DRL) with vision in simulation, sim-to-real with real flights, and sim-to-real methodology.

The thesis goal is a low-cost autonomous drone flying through a forest with one RGB camera ($62^{\circ}$ FOV), no LiDAR or stereo, depth recovered in software with Depth Anything. The platform is a full-size Holybro S500 V2, the algorithm is Proximal Policy Optimization (PPO), and training is in AirSim (indoor + outdoor scenes) before real tests in a Swedish forest.

\subsection{Overview and Research Gap}

Table \ref{tab:related_work} compares the closest studies along sensor, algorithm, test environment, and whether the drone actually flew.

\begin{table}[H]
\centering
\caption{Comparison of closely related studies along the axes used in this thesis.}
\label{tab:related_work}
\resizebox{\textwidth}{!}{%
\footnotesize
\begin{tabular}{lcccc}
\hline
\textbf{Study} & \textbf{Sensor} & \textbf{Algorithm} & \textbf{Test environment} & \textbf{Real flight} \\
\hline
Kumar et al.\ \cite{Kumar2023}        & Single RGB camera          & Classical CV                     & Indoor                          & Yes \\
Kabas \cite{Kabas2022}                & RGB or depth (AirSim)      & PPO                              & AirSim corridor with openings   & No  \\
Loquercio et al.\ \cite{Loquercio2019} & Monocular RGB (300$\times$200) & CNN + imitation + DR             & Racing track                    & Yes \\
Jang et al.\ \cite{Jang2024}          & Mixed                      & DRL + sim-to-real methods         & Mixed                           & Partial \\
Del Col et al.\ \cite{DelCol2025}     & RealSense D435i + T265     & Depth autoencoder + LSTM CPN     & Boreal Finnish forest           & Yes \\
\hline
\textbf{This thesis} & \textbf{Single RGB, $62^{\circ}$, + Depth Anything} & \textbf{PPO} & \textbf{AirSim + real Swedish forest} & \textbf{Yes} \\
\hline
\end{tabular}%
}
\end{table}

Three patterns stand out: deep-RL avoidance is usually only tested in simulation \cite{Kabas2022}; real-world studies use either racing drones on clean tracks \cite{Loquercio2019} or platforms with stereo and tracking cameras \cite{DelCol2025}; and richer sensor stacks work, but LiDAR (\$1{,}000--\$10{,}000) and stereo (\$300--\$800) are too expensive, heavy, and fiddly to calibrate for routine forest use \cite{Butt2024,Hrabar2012}. No earlier study combines PPO, a single narrow-FOV RGB camera with software depth from Depth Anything, a full-size quadrotor, AirSim training, and real forest flights -- the gap this thesis closes to remove the cost barrier to routine bark-beetle monitoring.

\subsection{Classical Single-Camera Navigation}

Software depth from a single RGB camera is now possible thanks to recent monocular depth models. Ranftl et al.\ \cite{Ranftl2022} showed with MiDaS that training across many depth datasets with a scale-invariant loss generalises to unseen scenes. Yang et al.\ \cite{Yang2024DepthAnythingV2} pushed this further with Depth Anything V2, trained on synthetic depth labels and pseudo-labelled real images, giving sharper depth and running $10\times$ faster than diffusion-based alternatives -- making software-only depth a real alternative to hardware sensors when cost is tight.

\paragraph{Kumar et al.\ \cite{Kumar2023} -- single-camera obstacle avoidance on a small onboard computer.}
Kumar et al.\ built a navigation pipeline using image-processing rules on a small onboard computer with a single RGB camera and MiDaS for depth -- direct support for the idea that one cheap camera can be enough. Weaknesses are hand-written rules rather than a learned policy and tidy indoor-only tests. This thesis replaces the rule-based pipeline with a CNN policy trained by PPO that learns directly from camera images, swaps MiDaS for Depth Anything V2, and tests outdoors against real tree trunks.

\subsection{DRL with Vision in Simulation}

\paragraph{Kabas \cite{Kabas2022} -- PPO with a single RGB or depth camera in AirSim corridors.}
Kabas trained a PPO agent end-to-end in an Unreal Engine + AirSim corridor with round openings. RGB alone reached 96\,\% success using a stack of grayscale frames through a Convolutional Neural Network (CNN) -- direct evidence that PPO + single camera suffices for non-trivial avoidance. Weaknesses are the regular corridor, minimal domain randomisation, and the policy never running outside simulation. This thesis swaps the corridor for irregular outdoor trees, borrows the software-depth idea from \cite{Kumar2023}, adds heavier domain randomisation in AirSim, keeps a CNN over a frame stack, and closes the loop with real flights.

\subsection{Sim-to-Real Studies with Real Flights}

\paragraph{Loquercio et al.\ \cite{Loquercio2019} -- Deep Drone Racing.}
A CNN reads a single $300\times200$ RGB image and outputs a local target direction and speed; a classical planner and controller produce motor commands. The network is trained by imitation learning on simulator-optimal trajectories with the scene randomised across background, lighting, gate shape, and gate texture. Versus a no-randomisation baseline, performance rose by 300\,\% in simulation and 36\,\% in the real world, showing zero-shot sim-to-real transfer for fast camera-based flight. An extraction ranks background randomisation essential, lighting second, gate-shape smallest -- evidence that domain randomisation alone can bridge sim and reality with one camera.

Weaknesses for forest use are a tidy track with bright gates, a small racer, and imitation of a pre-computed expert -- a recipe the authors say fails without a defined expert path. This thesis keeps the single-camera, randomised, sim-to-real philosophy but uses PPO instead of imitation learning, since a forest has no gates and no optimal path. Following the extraction, AirSim training randomises three forest analogues: trunk shape and angle (gate shape), time of day (lighting), and forest layout with randomised tree positions (background -- the most important factor). The track becomes an unstructured Swedish forest, the racer a Holybro S500 V2, and high-speed racing a safe forest cruise.

\paragraph{Del Col et al.\ \cite{DelCol2025} -- autonomous drone navigation in Finnish boreal forests.}
Del Col et al.\ (UAV-g 2025, Espoo) carry an Intel RealSense D435i stereo depth camera and an Intel RealSense T265 tracking camera, with onboard computation on an NVIDIA Orin NX and a Holybro Pixhawk 6C Mini. The action space is discrete: constant forward speed $\times$ 8 vertical-speed $\times$ 32 yaw-rate ($8\times32=256$ trajectories) scored each step. The Collision Prediction Network is trained by \emph{supervised} learning on simulator-generated collision labels (RotorS and Aerial Gym), not RL, and tested in real Finnish forests with autonomous flights around 60\,m in cluttered scenes -- the most directly comparable forest study. Weaknesses are the cost, weight, and calibration of the stereo and tracking cameras (\$300--\$800 \cite{Butt2024,Hrabar2012}).

The hardware/middleware base matches this thesis: the same Holybro/PX4 family (Pixhawk 6C vs.\ 6C Mini), the same NVIDIA Jetson Orin family (Orin Nano vs.\ Orin NX), and ROS as middleware. This thesis also uses discrete actions, but lets PPO learn which to pick directly from the camera image rather than scoring a fixed motion-primitive grid. The sensor side is deliberately cheaper and lighter: one monocular RGB camera with Depth Anything in software replaces the RealSense D435i + T265 pair.

\subsection{Sim-to-Real Methodology Studies}

\paragraph{Zhao et al.\ \cite{Zhao2020} and Salvato et al.\ \cite{Salvato2021} -- sim-to-real surveys.}
Two complementary surveys of sim-to-real transfer in deep RL review the standard toolkit -- domain randomisation, dynamics randomisation, system identification, and observation augmentation -- and single out sensor-noise injection (Gaussian image noise, brightness/contrast jitter, sensor delay, motor-command noise) as one of the cheapest and most reliable ways to harden a policy; neither offers a recipe specific to outdoor UAV flight. This thesis applies the conclusion directly: the sensor pipeline injects per-frame Gaussian image noise, brightness/contrast jitter, motion blur, randomised inference delay, and small action noise during training, tuned for the Depth Anything inference profile on the onboard Jetson.

\paragraph{Jang et al.\ \cite{Jang2024} -- deep-RL policies from simulation to reality.}
Jang et al.\ treat sim-to-real systematically across robotic platforms, concluding that careful domain randomisation plus manageable real-world tuning usually closes the gap. The case studies are not on outdoor forest flight -- which this thesis re-checks on a full-size drone with a single RGB camera.

No prior work combines PPO, a single narrow-FOV RGB camera with software depth from Depth Anything, asymmetric actor--critic training, and validation in a real forest. The methods chapter describes that combination, and the next subsection sets out the rationale behind each design choice.

\end{document}
