\documentclass[../main.tex]{subfiles}

\begin{document}

\section{Analysis and Discussions}
\justify

\subsection{Sparse vs.\ Dense Performance}

The agent consistently achieved higher success rates in sparse conditions than in dense conditions across both simulation and real-world trials. In simulation, episode-level success dropped from 94.00\% (sparse) to 84.00\% (dense), and in real-world trials from 74.00\% to 70.00\%. This pattern is expected: with 3 trees in the corridor instead of 2, the agent faces a higher probability of consecutive avoidance decisions with less lateral recovery space between them. A single failed avoidance causes a collision and ends the episode regardless of how well the agent handles all other trees.

The per-obstacle avoidance probabilities tell a complementary story. In dense real-world trials the estimated per-obstacle avoidance probability was 88.64\,\%, yet the episode success rate was only 70.00\,\%. This is consistent with the geometric model: $0.8864^3 \approx 0.696$, closely matching the observed episode success rate. The gap between per-obstacle and episode-level performance is therefore the expected cumulative effect of needing to avoid all trees in sequence, not a general degradation in avoidance ability. The agent's single-tree avoidance capability degrades less with density than its episode completion rate suggests.

\subsection{Sim-to-Real Transfer}

The overall episode-level transfer efficiency of 80.90\% and per-obstacle transfer of 91.76\% represent a strong zero-shot sim-to-real result, meaning the policy was deployed on the physical drone without any real-world fine-tuning or adaptation. This is consistent with the best results reported in the sim-to-real literature for vision-based drone navigation \cite{Loquercio2019,Jang2024,Azzam2024}.

We attribute this to the domain randomisation strategy applied during training. By randomising tree positions, angles, and trunk sizes across episodes, applying per-frame Gaussian image noise with brightness and contrast jitter, adding directional motion blur scaled by the lateral action, introducing camera-mounting jitter and spawn yaw jitter, injecting velocity drift of up to 0.3\,m/s, and randomising the inference pipeline delay between 100 and 200\,ms, the agent was exposed to a wide distribution of visual and physical conditions. This prevented overfitting to simulation-specific visual patterns and encouraged the policy to generalise to the real-world environment, where lighting, bark texture, and background vary in ways that cannot be fully anticipated in simulation.

Despite this, the remaining 19.10\% episode-level gap between simulation and real-world performance reflects residual sources of reality gap that domain randomisation did not fully bridge. These specific sources were not isolated experimentally; plausible candidates include aerodynamic differences between the simulated and physical Holybro S500 V2 platform, actuation latency between the moment a velocity command is sent to the PX4 flight controller and the moment the motors physically execute it, and minor camera calibration differences between the simulated sensor and the physical Arducam IMX219. Quantifying their individual contributions would require dedicated follow-up experiments that measure each factor in isolation, left to future work. The per-obstacle transfer rate of 91.76\% being considerably higher than the episode-level rate of 80.90\% suggests that the gap is not primarily a perception failure but a control precision issue: the agent sees the obstacles correctly but occasionally mistimes the avoidance manoeuvre in the real world due to physical dynamics the simulation did not perfectly capture.

\subsection{Dense Conditions Show Higher Transfer Efficiency}

A notable finding is that transfer efficiency was slightly higher for dense conditions (83.33\%) than for sparse (78.72\%) at the episode level. This is counterintuitive at first glance, since dense environments are harder. One plausible explanation is that the agent's learned policy in dense conditions relies more heavily on reactive, close-range avoidance behaviour that is less sensitive to small differences in flight dynamics between simulation and reality. In sparse conditions, the agent has more time and space to build up lateral velocity, meaning small differences in physical response can result in a wider trajectory deviation by the time the next tree is reached. In dense conditions, the shorter inter-tree spacing keeps the agent's corrections smaller and more frequent, making the policy more robust to individual control errors.

\subsection{Limitations Observed in Results}

The evaluation was conducted on cylindrical trunks under controlled conditions. The agent was not tested on branching trees, mixed species, or environments with ground cover that could occlude the lower trunk region. Performance in a genuine forest with greater visual complexity would likely be lower than reported here, and the current results should be interpreted as an upper bound for these controlled conditions rather than a direct estimate of deployment performance in operational bark beetle monitoring.

The dataset is also limited to sparse and dense configurations (2 and 3 trees in the corridor). A more complete evaluation would include a wider range of densities and longer corridors to better characterise how performance scales with environmental complexity.

The domain randomisation suite used in this thesis, while broad, leaves room for improvement. More aggressive randomisation (including dynamic obstacles, rain effects, varied lighting conditions, and a wider range of tree types and bark textures) could further close the sim-to-real gap. Whether the additional training complexity would yield proportional gains is an open question that motivates future work.

\end{document}
