\documentclass[../main.tex]{subfiles}

\begin{document}

\section{Conclusions}
\justify

Taken as a whole, the experimental results demonstrate considerable potential for the proposed approach. The simulated drone reached an episode-level success rate of 89.00\,\% (and a per-obstacle success rate of 95.60\,\%) across the two density conditions, and that performance carried over to the real forest with an episode-level transfer efficiency of 80.90\,\% and a per-obstacle transfer efficiency of 91.76\,\%. These results are statistically meaningful across 200 evaluation episodes and show that a domain-randomised PPO policy can transfer from simulation to a real forest without fine-tuning.

It is important to be clear about what these numbers do and do not imply for operational deployment. A per-obstacle avoidance probability of approximately 88\,\% in the real world means that, on average, roughly one in eight trees results in a collision. In a real forest survey -- where a drone might need to navigate past hundreds or thousands of trees -- this collision rate would make unsupervised autonomous operation impractical at current performance levels. The evaluation corridors in this thesis, containing 2--3 trees over 15\,m, are a necessary starting point for controlled measurement, but they represent a far simpler scenario than full forest coverage. The honest conclusion is therefore not that the problem is solved, but that the approach has demonstrated the potential to be a viable path forward.

The numbers also become more meaningful once the platform constraints are taken into account. The drone was equipped with only a single \$20 monocular camera, no LiDAR or stereo, on a full-size Holybro S500 V2 frame with a narrow $62^{\circ}$ field of view -- an unusually restrictive perception setup for forest flight. That this configuration already achieves above 80\,\% sim-to-real transfer efficiency suggests that the ceiling for this approach is considerably higher than what has been demonstrated here. Wider field of view, a lighter platform, and more aggressive domain randomisation are all directions that could plausibly improve the collision rate to levels where supervised autonomous surveys become practical.

\end{document}
